Note on: Insights, Intuition, and Connections regarding Laplacian Learning

### 2021-03-17 notes/misc/insights_and_connections.tex

\paragraph{Laplacian learning and RKHS: continuum limits when $d \geq 2$.}
\begin{itemize}
	\item When $d \geq 2$, the space $H_0^1(P)$ is no longer an RKHS.
	\item I claim that the right continuum counterpart to Laplacian smoothing is instead given by restricting the domain of minimization to the first $N$ eigenfunctions of $\Delta_P$. In other words, let $V_k = \mathrm{span}\{v_1,\ldots,v_k\}$, and take
	\begin{equation}
	\label{eqn:bandlimited_laplacian_smoothing}
	\check{u} := \argmin_{u \in V_N} \Bigl\{ \|y - u\|_n^2 + \lambda \int \|\nabla u\| p^2 \,d \nu  \Bigr\}
	\end{equation}
	The vector space $V_k$, equipped with inner product $\dotp{\cdot}{\cdot}_{\mc{H}}$, is an RKHS. Indeed, it is well-studied in the special case $d = 1$, $\Omega = \Reals$ and $P$ uniform. I further claim that as long as $N \to \infty$ sufficiently slowly as $n \to \infty$, the solution to~\eqref{eqn:bandlimited_laplacian_smoothing} is well-behaved in the limit.
	\item \textcolor{red}{(TODO):} We need to justify each of the claims in the prior point. 
	\begin{itemize}
		\item I believe the first claim can be rigorously justified when the domain is $[0,1]^d$, the design points are equi-spaced grid points, the graph is the lattice, and the measure $P$ is uniform. It is much trickier to justify when the graph is a neighborhood graph, because it involves making statements about the top eigenvectors of the Laplacian $L_{N,\varepsilon}$, which are poorly understood. If we cannot rigorously justify it in this case, then the next best option is to give some intuition. For instance, we could say that all eigenvectors of $L_{N,\varepsilon}$ can be extrapolated to locally Lipschitz functions on $\Omega$ with bounded Lipschitz constant, and that such functions in turn lie very close to $V_N$.
		\item The second claim is simply a matter of adapting some standard results regarding the entropy of RKHS.
	\end{itemize}
	\item \textcolor{red}{(TODO)}: We should also give some insight into the structure of $V_N$. For instance, what are the smoothness properties of functions $u \in V_N$? What does the kernel function look like (this is well known, I believe, when $d = 1$ and $p$ is uniform)? How do these depend on the density $p$?
	\item \textcolor{red}{(TODO)}: Some grumpy advisor might say that we need to justify why anyone should care about relating $\wh{u}$ to $\check{u}$. One way to do this is by arguing that methodologically~\eqref{eqn:bandlimited_laplacian_smoothing} is superior to~\eqref{eqn:laplacian_smoothing}, for instance because it is well defined over all of $\Omega$ rather than only $\Omega_N$. Another way to do this is by arguing that the representation~\eqref{eqn:bandlimited_laplacian_smoothing} buys us something theoretically. Three example things it might buy are: 
	\begin{enumerate}[(1)]
		\item It may be possible to relate~\eqref{eqn:bandlimited_laplacian_smoothing} to an equivalent kernel,
		\item It is far easier to analyze~\eqref{eqn:bandlimited_laplacian_smoothing} than directly analyzing Laplacian smoothing. Hopefully, we can easily get sharp upper bounds for more interesting situations than we have analyzed thus far. For instance, when $p$ or $\Omega$ are complicated, instead of being a smooth function with a smooth boundary, respectively. 
		\item It should be easy to prove a lower bound on the error of~\eqref{eqn:bandlimited_laplacian_smoothing} when $d > 4$. 
	\end{enumerate}
\end{itemize}

>>>>

\paragraph{Laplacian learning and RKHS: continuum limits when $d \geq 2$.}
\begin{itemize}
	\item When $d \geq 2$, the space $H_0^1(P)$ is no longer an RKHS.
	\item I claim that the right continuum counterpart to Laplacian smoothing is instead given by restricting the domain of minimization to the first $N$ eigenfunctions of $\Delta_P$. In other words, let $V_k = \mathrm{span}\{v_1,\ldots,v_k\}$, and take
	\begin{equation}
	\label{eqn:bandlimited_laplacian_smoothing}
	\check{u} := \argmin_{u \in V_N} \Bigl\{ \|y - u\|_n^2 + \lambda \int \|\nabla u\| p^2 \,d \nu  \Bigr\}
	\end{equation}
	The vector space $V_k$, equipped with inner product $\dotp{\cdot}{\cdot}_{\mc{H}}$, is an RKHS. Indeed, it is well-studied in the special case $d = 1$, $\Omega = \Reals$ and $P$ uniform. I further claim that as long as $N \to \infty$ sufficiently slowly as $n \to \infty$, the solution to~\eqref{eqn:bandlimited_laplacian_smoothing} is well-behaved in the limit.
	\item \textcolor{red}{(TODO):} We need to justify each of the claims in the prior point. 
	\begin{itemize}
		\item I believe the first claim can be rigorously justified when the domain is $[0,1]^d$, the design points are equi-spaced grid points, the graph is the lattice, and the measure $P$ is uniform. It is much trickier to justify when the graph is a neighborhood graph, because it involves making statements about the top eigenvectors of the Laplacian $L_{N,\varepsilon}$, which are poorly understood. If we cannot rigorously justify it in this case, then the next best option is to give some intuition. For instance, we could say that all eigenvectors of $L_{N,\varepsilon}$ can be extrapolated to locally Lipschitz functions on $\Omega$ with bounded Lipschitz constant, and that such functions in turn lie very close to $V_N$.
		\item The second claim is simply a matter of adapting some standard results regarding the entropy of RKHS.
	\end{itemize}
	\item \textcolor{red}{(TODO)}: We should also give some insight into the structure of $V_N$. For instance, what are the smoothness properties of functions $u \in V_N$? What does the kernel function look like (this is well known, I believe, when $d = 1$ and $p$ is uniform)? How do these depend on the density $p$?
	\item \textcolor{red}{(TODO)}: Some grumpy advisor might say that we need to justify why anyone should care about relating $\wh{u}$ to $\check{u}$. One way to do this is by arguing that methodologically~\eqref{eqn:bandlimited_laplacian_smoothing} is superior to~\eqref{eqn:laplacian_smoothing}, for instance because it is well defined over all of $\Omega$ rather than only $\Omega_N$. Another way to do this is by arguing that the representation~\eqref{eqn:bandlimited_laplacian_smoothing} buys us something theoretically. Three example things it might buy are: 
	\begin{enumerate}[(1)]
		\item It may be possible to relate~\eqref{eqn:bandlimited_laplacian_smoothing} to an equivalent kernel,
		\item It is far easier to analyze~\eqref{eqn:bandlimited_laplacian_smoothing} than directly analyzing Laplacian smoothing. Hopefully, we can easily get sharp upper bounds for more interesting situations than we have analyzed thus far. For instance, when $p$ or $\Omega$ are complicated, instead of being a smooth function with a smooth boundary, respectively. 
		\item It should be easy to prove a lower bound on the error of~\eqref{eqn:bandlimited_laplacian_smoothing} when $d > 4$. 
	\end{enumerate}
	In a negative direction,~\eqref{eqn:bandlimited_laplacian_smoothing} reveals that the Laplacian smoothing is automatically zeroing out certain high-frequency signals. Once we acknowledge this fact, we should probably also acknowledge that the particular choice of frequency at which to cut off was made somewhat arbitrarily. Indeed, we get much stronger theoretical guarantees if we cut off at a much lower frequency. This motivates Laplacian eigenmaps.
\end{itemize}
